Author disambiguation of large-scale semi-labeled collections of scientific articles

### 2015-08-05 paper.tex
save

\title{Semi-supervised author disambiguation of large-scale collections of labeled scientific publications}

@@

Efforts and solutions to author disambiguation have been proposed from various
communities \citep{liu2014author}. On the one hand, libraries have maintained
authorship control through manual curation, either in a centralized way by
hiring collaborators or by developing services inviting authors to register
their publications themselves (e.g., Google Scholar, Inspire-HEP, ...). Recently,
efforts also started to create persistent digital identifiers assigned to researchers
(e.g., ORCID, ResearchedID, ...), with the further objective to embed these
identifiers in the submission workflow of publishers or repositeries (e.g.,
Elsevier, arXiv, Inspire-HEP, ...), thereby univocally solving any disambiguation
issue. With the large cost of centralized manual authorship control or until
the larger adoption of crowdsourced solutions, the impact of these efforts are
unfortunately limited by the efficiency, motivation or integrity of their
active contributors. Similarly, the success of persistent digital identifier
efforts is conditioned to a large and ubiquituous adoption by both researchers and publishers.
For these reasons, fully automated machine learning-based methods have been
proposed during the past decade (See Section~\ref{related-works}) to provide immediate, less costly
and efficient solutions to author disambiguation. In this work, our goal
is to explore and demonstrate how both approaches can coexist and benefit from
each other.  In particular, we study how labeled data obtained through manual
curation (either centralized or crowdsourced) can be exploited (i) to learn an
accurate classifier for identifying corefering authors, (ii) to guide the
clustering of scientific publications by distinct authors in a semi-supervised
way and finally (iii) to match the resulting clusters with existing profiles.
Our analysis of parameters and features on this large training set reveal
that... \glnote{To be completed with actual results.}

@@

As reviewed in
\citep{smalheiser2009author,ferreira2012brief,levin2012citation}, author
disambiguation algorithms are usually composed of two components: (i) a
similarity function determining whether two publications have been written by a
same author and (ii) a clustering algorithm producing clusters of publications
assumed to be written by a same author. Approaches can be inventorized along
several axes, depending on the type and amount of data available, the way the
similarity function is learned or defined, or the clustering procedure used to
group publications. Methods relying on supervised learning usually make use of
hand-labeled pairs of publications identified as being either from the same or
from different authors for learning automatically a similarity function between
publications \citep{han2004two,huang2006efficient,
culotta2007author,treeratpituk2009disambiguating,tran2014author}. Because
training data is often not easily available, unsupervised approaches propose
instead to use domain-specific similarity functions designed for author
disambiguation
\citep{malin2005unsupervised,mcrae2006also,song2007efficient,kang2009co,fan2011graph,schulz2014exploiting}.
These later approaches have the advantage of requiring no hand-labeled data, but
often do not perform as well as the supervised approaches. To reconcile
both worlds, semi-supervised

@@

In this work, author disambiguation is cast into a semi-supervised clustering
problem, as inspired from several previous works \glnote{Add references}.
Our algorithm is made of four elements: ...

>>>>

\title{Author disambiguation of large-scale semi-labeled collections of scientific articles}

@@

Efforts and solutions to author disambiguation have been proposed from various
communities \citep{liu2014author}. On the one hand, libraries have maintained
authorship control through manual curation, either in a centralized way by
hiring collaborators or by developing services inviting authors to register
their publications themselves (e.g., Google Scholar, Inspire-HEP, ...). Recently,
efforts also started to create persistent digital identifiers assigned to researchers
(e.g., ORCID, ResearchedID, ...), with the further objective to embed these
identifiers in the submission workflow of publishers or repositeries (e.g.,
Elsevier, arXiv, Inspire-HEP, ...), thereby univocally solving any disambiguation
issue. With the large cost of centralized manual authorship control or until
the larger adoption of crowdsourced solutions, the impact of these efforts are
unfortunately limited by the efficiency, motivation or integrity of their
active contributors. Similarly, the success of persistent digital identifier
efforts is conditioned to a large and ubiquituous adoption by both researchers and publishers.
For these reasons, fully automated machine learning-based methods have been
proposed during the past decade to provide immediate, less costly
and efficient solutions to author disambiguation. In this work, our goal
is to explore and demonstrate how both approaches can coexist and benefit from
each other.  In particular, we study how labeled data obtained through manual
curation (either centralized or crowdsourced) can be exploited (i) to learn an
accurate classifier for identifying corefering authors, (ii) to guide the
clustering of scientific publications by distinct authors in a semi-supervised
way and finally (iii) to match the resulting clusters with existing profiles.
Our analysis of parameters and features on this large training set reveal
that... \glnote{To be completed with actual results.}

@@

As reviewed in
\citep{smalheiser2009author,ferreira2012brief,levin2012citation}, author
disambiguation algorithms are usually composed of two components: (i) a
similarity function determining whether two publications have been written by a
same author and (ii) a clustering algorithm producing clusters of publications
assumed to be written by a same author. Approaches can be inventorized along
several axes, depending on the type and amount of data available, the way the
similarity function is learned or defined, or the clustering procedure used to
group publications. Methods relying on supervised learning usually make use of
hand-labeled pairs of publications identified as being either from the same or
from different authors for learning automatically a similarity function between
publications \citep{han2004two,huang2006efficient,
culotta2007author,treeratpituk2009disambiguating,tran2014author}. Because
training data is often not easily available, unsupervised approaches propose
instead to use domain-specific similarity functions tailored for author
disambiguation \citep{malin2005unsupervised,mcrae2006also,song2007efficient,
soler2007separating, kang2009co,fan2011graph,schulz2014exploiting}. These later
approaches have the advantage of not requiring  hand-labeled data, but often do
not perform as well as the supervised approaches. To reconcile both worlds,
semi-supervised methods make use of small, manually verified, clusters of
publications and/or of high-precision domain-specific rules to build a training
set of pairs of publications, from which a similarity function is then built
using supervised learning
\citep{ferreira2010effective,torvik2009author,levin2012citation}.
Coincidentally, semi-supervised approaches also allow for the tuning of the
clustering algorithm when the later is applied to a mixed set of labeled
and unlabeled publications, e.g., by maximizing some clustering performance
metric on the known clusters \citep{levin2012citation}.

Because of the lack of a large and publicly available dataset of curated
clusters of publications, studies on author disambiguation are usually
constrained to validate their results on manually built datasets of limited
size and scope (from a few hundreds to a few thousands of papers, with a sparse
coverage of ambiguous cases), making the true performance of these methods
often difficult to assess with high confidence. In addition, despite devoted
efforts to construct them, these datasets are very rarely released publicly,
making it even more difficult to compare methods on a common benchmark.

In this context, we position the work presented in this paper as a
semi-supervised  solution for author disambiguation, but with the significant
advantage of having a very large collection of more than 1 million annotations
of publications known to be from the same author. In particular, the extent and coverage
of this data allows us to revisit, validate and nuance previous findings regarding
supervised learning of similarity functions.

@@

In this work, author disambiguation is cast into a semi-supervised clustering
problem, as inspired from several previous works described in Section~\ref{related-works}.
Our algorithm is made of four elements: ...
